Reanalysis Step 4: Main Effects Outcome Analysis

### 2025-08-28 reanalysis-4-outcome_analysis.qmd
accept

# Get number of imputations from saved data
n_imputations <- length(imputed_datasets)

@@

**Analysis Parameters:**
- Mode: `r analysis_mode`
- Number of imputations: `r n_imputations`
- Estimand: ATT (Average Treatment Effect in the Treated)
- Model: Main effects only (no interactions)

@@

summary_df <- data.frame(
  Measure = c("Total patients", "Lapsed", "Not lapsed", 
              "VI/Blind outcomes", "VI/Blind proportion"),
  Value = c(n_total, n_lapsed, n_control, n_vi, paste0(round(prop_vi, 1), "%"))
)

@@

**Treatment Status:**
```{r treatment-dist}
treatment_table <- table(first_imp$any_treatment)
treatment_df <- data.frame(
  Treatment = c("No Treatment", "Treatment"),
  Count = as.numeric(treatment_table),
  Percentage = paste0(round(100 * as.numeric(treatment_table) / sum(treatment_table), 1), "%")
)
kable(treatment_df, align = c('l', 'r', 'r'))
```

@@

```{r define-formula}
# Base covariates (same as interaction model)
base_covariates <- c(
  "baseline_VA_logMAR",
  "gender_cat",
  "race_ethnic_cat", 
  "insurance_cat",
  "age_cat",
  "CCI",
  "DCSI",
  "glaucoma_bef_hitplus_cat",
  "otherretina_bef_hitplus_cat",
  "catsurg_before_hitplus_cat"
)

# Main effects formula - NO INTERACTIONS
# Lapse is the exposure, DR severity and treatment are covariates
main_effects_formula <- as.formula(paste(
  "outcome_va_vi_binary ~ ever_lapse_binary + dr_severity + any_treatment +",
  paste(base_covariates, collapse = " + ")
))
```

@@

# Create diagnostics table
pool_diag <- data.frame(
  Metric = c("Number of imputations", 
             "Number of parameters",
             "Average degrees of freedom",
             "Fraction of missing information (FMI)",
             "Relative efficiency"),
  Value = c(n_imputations,
            nrow(pooled_results),
            ifelse("df" %in% names(pooled_results), round(mean(pooled_results$df, na.rm = TRUE), 1), NA),
            ifelse(!is.na(fmi_value), round(fmi_value, 3), "Not available"),
            ifelse(!is.na(fmi_value), round(1 / (1 + fmi_value/n_imputations), 4), "Not available"))
)

>>>>

# Determine number of imputations to use based on analysis mode
n_imputations_available <- length(imputed_datasets)
if (analysis_mode == "test") {
  n_imputations <- min(2, n_imputations_available)  # Use only 2 for testing
  # Subset to first 2 imputations for testing
  imputed_datasets <- imputed_datasets[1:n_imputations]
  twang_weights <- twang_weights[1:n_imputations]
} else {
  n_imputations <- n_imputations_available  # Use all available
}

@@

**Analysis Parameters:**
- Mode: `r analysis_mode`
- Number of imputations: `r n_imputations` `r if(analysis_mode == "test") paste0("(", n_imputations_available, " available, using first ", n_imputations, " for testing)") else ""`
- Estimand: ATT (Average Treatment Effect in the Treated)

@@

summary_df <- data.frame(
  Measure = c("Total patients", "Lapsed", "Not lapsed", 
              "VI outcomes", "VI proportion"),
  Value = c(n_total, n_lapsed, n_control, n_vi, paste0(round(prop_vi, 1), "%"))
)

@@

@@

```{r define-formula}
# Load the standard formulas from functions.R
source("functions.R")

# Get the standard analysis formulas used in the original analysis
standard_formulas <- get_analysis_formulas()

# For reference, the original full formula is:
# standard_formulas$full

# For the reanalysis, we extend the full formula by adding:
# - dr_severity: diabetic retinopathy severity (No_DR, NPDR, PDR)
# - any_treatment: whether patient received any DR treatment
# These are new covariates not in the original analysis
main_effects_formula <- as.formula("outcome_va_vi_binary ~ 
  ever_lapse_binary + 
  dr_severity +
  any_treatment +
  baseline_VA_logMAR +
  gender_cat +
  race_ethnic_cat +
  insurance_cat +
  age_cat +
  CCI +
  DCSI +
  other_inject +
  anti_VEGF +
  focal_laser_flag +
  PRP_flag +
  glaucoma_bef_hitplus_cat +
  otherretina_bef_hitplus_cat +
  catsurg_before_hitplus_cat")
```

@@

# Create diagnostics table
pool_diag <- data.frame(
  Metric = c("Number of imputations", 
             "Number of parameters",
             "Average degrees of freedom"),
  Value = c(n_imputations,
            nrow(pooled_results),
            ifelse("df" %in% names(pooled_results),
             round(mean(pooled_results$df, na.rm = TRUE), 1), NA))
)
